---
title: "One-way ANOVA"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
  html_notebook:
    toc: yes
editor_options:
  chunk_output_type: inline
---
The ANOVA test is based on the cleaned data from regression step, dedicated to show the association between variables and the outcome.

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(viridis)
library(plotly)
library(modelr)
library(gtsummary)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 6, 
  fig.height = 8,
  out.width = "90%",
	message = F,
  echo = T,
  warning = F,
  cache = F
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```


```{r data import, echo=FALSE, message=FALSE, warning=FALSE}
nyc_shooting = read_csv("./comp_data/Full NYC Shooting Data (2006-2021).csv")

nyc_time = 
  nyc_shooting %>% 
  separate(occur_time, into = c("hour", "minute", "second"), sep = ":") %>% 
  mutate(
    hour = as.factor(hour)
  )
```

```{r, import data, echo=FALSE, message=FALSE, warning=FALSE}
nyc_fit_lm =
  nyc_time %>% 
  select(month, day, year, boro, statistical_murder_flag, perp_age_group, vic_age_group, perp_sex, perp_race, vic_sex, vic_race) %>%
  mutate(
    month = as.factor(month), 
    day = as.factor(day), 
    year = as.factor(year),
    boro = as.factor(boro), 
    statistical_murder_flag = as.factor(statistical_murder_flag),
    perp_sex = as.factor(perp_sex),
    perp_race = as.factor(perp_race),
    vic_sex = as.factor(vic_sex),
    vic_race = as.factor(vic_race),
    month = fct_relevel(month, "January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
  ) %>% 
  group_by(month, year, boro, statistical_murder_flag, perp_age_group, vic_age_group, perp_sex, perp_race, vic_sex, vic_race) %>% 
  summarize(
    number_shoot = n()
  ) %>% 
  drop_na()
```

```{r function showing}
showing = function(x){
  x %>% 
  broom::tidy() %>%
  mutate(
    p.value = format(p.value, scientific = TRUE, digits = 3)
  ) %>% 
  select(term, p.value) %>% 
  rows_delete(tibble(term = "Residuals")) %>% 
  rename(variable = term) %>% 
  knitr::kable()
}
```


# Year
```{r year}
res_year = aov(number_shoot ~ factor(year), data = nyc_fit_lm)
showing(res_year)
```

The p-value of the year variable is smaller than 0.05, thus we reject the null hypothesis and regard the year as statistically significant.

# Month

```{r month}
res_month = aov(number_shoot ~ factor(month), data = nyc_fit_lm)
showing(res_month)
```
The p-value of the month variable is smaller than 0.05, thus we reject the null hypothesis and regard the month as statistically significant.

# Borough

```{r boro}
nyc_boro = nyc_fit_lm %>% 
  rename(borough = boro)
res_boro = aov(number_shoot ~ factor(borough), data = nyc_boro)
showing(res_boro)
```

The p-value of the borough variable is smaller than 0.05, thus we reject the null hypothesis and regard the borough as statistically significant.

# Statistical_murder_flag

```{r Statistical_murder_flag}
res_smf = aov(number_shoot ~ factor(statistical_murder_flag), data = nyc_fit_lm)
showing(res_smf)
```

The p-value of the Statistical_murder_flag variable is smaller than 0.05, thus we reject the null hypothesis and regard the Statistical_murder_flag as statistically significant.

# Perp_sex

```{r perp_sex}
res_perpsex = aov(number_shoot ~ factor(perp_sex), data = nyc_fit_lm)
showing(res_perpsex)
```

The p-value of the sex of perpetrator variable is smaller than 0.05, thus we reject the null hypothesis and regard the sex of perpetrator as statistically significant.

# Perp_race

```{r perp_race}
res_perprace = aov(number_shoot ~ factor(perp_race), data = nyc_fit_lm)
showing(res_perprace)
```

The p-value of the race of perpetrator variable is smaller than 0.05, thus we reject the null hypothesis and regard the race of perpetrator as statistically significant.

# Vic_sex

```{r vit_sex}
res_vicsex = aov(number_shoot ~ factor(vic_sex), data = nyc_fit_lm)
showing(res_vicsex)
```

The p-value of the sex of victims variable is smaller than 0.05, thus we reject the null hypothesis and regard the sex of victims as statistically significant.

# Vic_race

```{r vic_race}
res_vicrace = aov(number_shoot ~ factor(vic_race), data = nyc_fit_lm)
showing(res_vicrace)
```

The p-value of the race of victims variable is smaller than 0.05, thus we reject the null hypothesis and regard the race of victims as statistically significant.

# Perp_age_group

```{r Perp_age_group}
res_perp_age_group = aov(number_shoot ~ factor(perp_age_group), data = nyc_fit_lm)
showing(res_perp_age_group)
```

The p-value of the perp_age_group variable is smaller than 0.05, thus we reject the null hypothesis and regard the perp_age_group as statistically significant.

# Vic_age_group

```{r Vic_age_group}
res_vic_age_group = aov(number_shoot ~ factor(vic_age_group), data = nyc_fit_lm)
showing(res_vic_age_group)
```

The p-value of the vic_age_group variable is smaller than 0.05, thus we reject the null hypothesis and regard the vic_age_group as statistically significant.
